\section{Collections}

Traditional user interfaces present the user with a window for each task
the user needs to accomplish.  Since we don't have the option of
presenting multiple "windows" to users of a terminal interface, we must
present the user with one interface at a time.  Then, the application
will manage the transition between these interfaces.

Consider a text editor program in which we must present these top-level
interfaces in the following order:

\begin{itemize}
\item The user runs the program and is presented with an interface to
      select a file to edit
\item The user chooses a file to edit and is presented with the
      editing interface
\item After editing, the user chooses to exit and we present a dialog
      which asks the user whether to save the file
\end{itemize}

All three of these interfaces are separate and should be given the
entire terminal window; unlike other graphical toolkits, \vtyui\ does
not provide a way to "show" or "hide" widgets.  Instead, it provides
the notion of a "collection."  A collection is a widget which wraps a
set of other widgets and maintains a pointer to the one that should be
displayed at any given time.  The application then changes the current
interface by changing the collection state.

But an interface is more than what is presented in the terminal; each
interface should have its own set of user input widgets and its own
notion of focus.  Therefore, a collection is a set of interfaces *and
their focus groups.*  When we change the state of the collection, we are
really changing both the visual interface as well as the focus group
used to interact with it.

To create a collection:

\begin{haskellcode}
 c <- newCollection
\end{haskellcode}

To add an interface and a focus group to the collection:

\begin{haskellcode}
 fg <- newFocusGroup
 -- Add widgets to focus group fg
 ui <- someWidget
 changeToW <- addToCollection c ui fg
\end{haskellcode}

As a convenience, addToCollection returns a MonadIO action which, when
run, will switch to the specified interface.  In the example above,
changeToW is an action which will switch to the interface with \fw{ui}
as its top-level widget and \fw{fg} as its focus group.  You can use
this action in event handlers that change your interface state.  If
you prefer, you can use the \fw{setCurrentEntry} function instead,
which allows you to set the collection's interface by number.  That
method is not recommended, however, since a bad index can cause an
exception to be thrown.
