\section{Fixed-Size Widgets}
\label{sec:fixed}

The \fw{Fixed} module provides widget containers which fix the amount
of spaced used to render the child.  This can be useful when you know
that an element of your interface has the potential to fill available
space but must be fixed to a specific size for some reason.

The module provides widget types for constraining the horizontal or
vertical size of a widget.  The fixed-size widget containers are
created with the following functions:

\begin{itemize}
\item \fw{hFixed} -- takes a widget \fw{Widget a} and a width in
  columns and constrains the widget to the specified width.  Returns a
  widget of type \fw{Widget (HFixed a)}.  If the \fw{HFixed} widget
  does not have enough space to enforce the specified width, the
  available space is used instead.
\item \fw{vFixed} -- takes a widget \fw{Widget a} and a height in rows
  and constrains the widget to the specified height.  Returns a widget
  of type \fw{Widget (VFixed a)}.  If the \fw{VFixed} widget does not
  have enough space to enforce the specified height, the available
  space is used instead.
\item \fw{boxFixed} -- takes a widget \fw{Widget a}, a width in
  columns, and a height in rows and constrains the widget in both
  dimensions.  Returns a widget of type \fw{Widget (VFixed (HFixed
    a))}.
\end{itemize}

In addition to widget creation, some manipulation functions are
provided so that the fixed-size container settings can be manipulated
as desired:

\begin{itemize}
\item \fw{setVFixed}, \fw{setHFixed} -- sets the constraint value for
  a fixed-size widget.
\item \fw{addToVFixed}, \fw{addToHFixed} -- adds a value to the
  constraint value of a fixed-size widget.
\item {getVFixedSize}, \fw{getHFixedSize} -- returns the constraint
  value of a fixed-size widget.
\end{itemize}

For example, the \fw{List} widget type (Section XXX) grows vertically,
but we may wish to dedicate most of the terminal to the rest of the
interface.  We can use \fw{vFixed} to constrain the \fw{List} in this
way.  Below, we constain the \fw{List} to five rows of space.
Assuming the \fw{List} elements are one row high each, if the
\fw{List} has fewer than five elements to display then the \fw{VFixed}
widget will automatically pad the \fw{List} to ensure that it takes up
the specified number of rows.  Fixed-size widgets thus provide both
upper and lower bounds on the size of their children.

\begin{haskellcode}
 lst <- newList (green `on` black) plainText
 ui <- vFixed 5 lst
\end{haskellcode}

\subsection{Growth Policy}

Since \fw{VFixed} and \fw{HFixed} widgets are designed to constrain
their children in a specific dimension, they never grow in the
constrained dimension.  For the other dimension, fixed-size widgets
always defer to their children for the growth policy.
