\section{Limits}
\label{sec:limits}

The \fw{Limits} module provides widgets for setting upper bounds on
the size of other widgets.  These widgets differ from the \fw{Fixed}
module we saw in Section \ref{sec:fixed}; ``limit'' widgets do not pad
their children if the children render to \fw{Image}s smaller than the
specified space, whereas fixed-size widgets pad their children in,
thus guaranteeing that the specified space will be consumed.

The limiting widget API is similar to that of the \fw{Fixed} module.
Limiting widgets are created as follows:

\begin{itemize}
\item \fw{hLimit} -- takes a widget \fw{Widget a} and a width in
  columns and constrains the widget to the specified width.  Returns a
  widget of type \fw{Widget (HLimit a)}.  If the \fw{HLimit} widget
  does not have enough space to enforce the specified width, the child
  widget is not padded.
\item \fw{vLimit} -- takes a widget \fw{Widget a} and a height in rows
  and constrains the widget to the specified height.  Returns a widget
  of type \fw{Widget (VLimit a)}.  If the \fw{VLimit} widget does not
  have enough space to enforce the specified height, the child widget
  is not padded.
\item \fw{boxLimit} -- takes a widget \fw{Widget a}, a width in
  columns, and a height in rows and constrains the widget in both
  dimensions.  Returns a widget of type \fw{Widget (VLimit (HLimit
    a))}.  If the child widget is smaller, it is not padded.
\end{itemize}

In addition to widget creation, some manipulation functions are
provided so that the limit settings can be manipulated as desired:

\begin{itemize}
\item \fw{setVLimit}, \fw{setHLimit} -- sets the constraint value for
  a limiting widget.
\item \fw{addToVLimit}, \fw{addToHLimit} -- adds a value to the
  constraint value of a limiting widget.
\item {getVLimitSize}, \fw{getHLimitSize} -- returns the constraint
  value of a limiting widget.
\end{itemize}

\subsection{Growth Policy}

Limiting widgets never grow in the constrained dimension, and defer to
their children for growth policy otherwise.
