\section{Padding}
\label{sec:padding}

The \fw{Padding} module provides a wrapper widget type, \fw{Padded},
which wraps another widget with a specified amount of padding on any
or all four of its sides.

We create padded widgets with the \fw{padded} function, which takes a
child of type \fw{Widget a} and a padding value.  In the following
example we create a \fw{FormattedText} widget and pad it on all sides
by two rows (or columns, where appropriate):

\begin{haskellcode}
 w <- plainText "foobar"
 w2 <- padded w (padAll 2)
\end{haskellcode}

The padding itself is expressed with the \fw{Padding} type, whose
values store padding settings for the top, bottom, left, and right
sides of an object in question.  \fw{Padding} values are created with
one of the following functions:

\begin{itemize}
\item \fw{padNone} -- creates a \fw{Padding} value with no padding.
\item \fw{padAll} -- takes a single parameter, \fw{p}, and creates a
  \fw{Padding} value with \fw{p} rows or columns of padding on all
  four sides.
\item \fw{padLeft}, \fw{padRight}, \fw{padTop}, \fw{padBottom} -- each
  takes a single parameter and creates a \fw{Padding} value with the
  specified amount of padding on the specified side indicated by the
  function name.
\item \fw{padLeftRight}, \fw{padTopBottom} -- each takes a single
  parameter and creates a \fw{Pad\-ding} value with the specified
  amount of padding on both sides indicated by the function name.
\end{itemize}

With these basic \fw{Padding} constructors we can construct more
interesting \fw{Padding} values with the \fw{pad} function:

\begin{haskellcode}
 let p = padNone `pad` (padAll 5) `pad` (padLeft 2)
\end{haskellcode}

The \fw{Padding} type is an instance of the \fw{Paddable} type class,
of which \fw{pad} is the only method.  The \fw{Padding} instance of
\fw{Paddable} just adds the padding values together.

In addition to the \fw{padded} function, the \fw{Padding} module
provides the \fw{withPadding} combinator to created a \fw{Padded}
widget in the following way:

\begin{haskellcode}
 w <- plainText "foobar" >>= withPadding (padAll 2)
\end{haskellcode}

\subsubsection{Growth Policy}

\fw{Padded} widgets always defer to their children for both horizontal
and vertical growth policy.
